comgen-outline

### 2017-12-22 paper/case1-generics.Rmd
minor edits; add footnote about hurdle models

<!-- Accurately modeling the prior data is important to generate precise quantitative predictions about endorsement. -->
In order to integrate the results of this prior measurement task into our model of endorsement, we must accurately model the prior elicitation data.
We do this using a Bayesian mixture model, which assumes that the data generated for each property (e.g., \textsc{lay eggs}) comes from one of two underlying Beta distributions (*Mixture of Betas*).^[
  The Beta distribution is chosen because the support of this distribution is numerical values between 0 - 1, exactly the form of the response data in the prior elicitation task.
] 
We specify one of these distributions *a priori* to represent kinds of animals who *do not have* a stable causal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}), which results in prevalence or prevalence values close to or equal to 0.
This "null distribution" is potentially present for all features in exactly the same way.
The second distribution represents kinds of animals who *do have* such a mechanism, and the two parameters of this distribution are not specified *a priori* and are not the same for all properties, but are inferred on a property-wise basis from participants' responses.
The *Mixture of Betas* distribution has a third free parameter (for each property), the relative contribution of the "null distribution" (for example: we expect the "null distribution" to not contribute at all to properties like *being female*, for which almost all categories have at least some members with the property).
To ensure the *Mixture of Betas* model of the prior is not overly complex, we fit an additional model that represents only a single underlying distribution (*Single Beta*) for a comparison.
<!-- is flexible and can represent different shapes for different properties (e.g., the differences in prevalence judgments for \textsc{lay eggs} vs. \text{has wings}). -->
For more details about model implementation and inference, see Appendix C.

>>>>

<!-- Accurately modeling the prior data is important to generate precise quantitative predictions about endorsement. -->
In order to integrate the results of this prior measurement task into our model of endorsement, we must accurately model the prior elicitation data.
We do this using a Bayesian mixture model, which assumes that the data generated for each property (e.g., \textsc{lay eggs}) comes from one of two underlying Beta distributions (*Mixture of Betas*).^[
  The Beta distribution is chosen because the support of this distribution is numerical values between 0 - 1, exactly the form of the response data in the prior elicitation task.
] 
We specify one of these distributions *a priori* to represent kinds of animals who *do not have* a stable causal mechanism that could give rise to the property (e.g., \textsc{lions} and \textsc{lay eggs}), which results in prevalence or prevalence values close to or equal to 0.^[
  This assumption is similar in spirit that employed by *Hurdle Models* of epidemiological data, where the observed count of zeros is often substantially greater than one would expect from standard models, such as the Poisson [e.g., when modeling adverse reactions to vaccines; @hurdleModels]
]
This "null distribution" is potentially present for all features in exactly the same way (i.e., the lack of producing the feature).
The second distribution represents kinds of animals who *do have* such a mechanism, and the two parameters of this distribution are not specified *a priori* and are not the same for all properties, but are inferred on a property-wise basis from participants' responses.
The *Mixture of Betas* distribution has a third free parameter (for each property), the relative contribution of the "null distribution" (for example: we expect the "null distribution" to not contribute at all to properties like *being female*, for which almost all categories have at least some members with the property).
To ensure the *Mixture of Betas* model of the prior is not overly complex, we fit an additional model that represents only a single underlying distribution (*Single Beta*) for a comparison.
<!-- is flexible and can represent different shapes for different properties (e.g., the differences in prevalence judgments for \textsc{lay eggs} vs. \text{has wings}). -->
For more details about model implementation and inference, see Appendix C.

### 2017-12-22 paper/case2-habituals.Rmd
minor edits; add footnote about hurdle models

```{r habituals-endorsement-gender-exploration-scatter, fig.cap='Exploratory analysis of endorsements of habitual statements by gender of target character (e.g., "Mary drinks beer" vs. "John...") for six events that displayed appreciable differences in the prior elicitation. Labeled points all correspond to the time period of 3 times / month.', eval=F}
gendered.items <- c("does cocaine", "drinks beer", "drinks coffee",
                    "wears a bra", "wears a watch", "wears a suit")

@@

hab.uncertain.threshold.s1opt <- m.hab.samp %>% 
  filter(type == "param") %>%
  group_by(param) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

habituals.endorsement.models <- md.hab %>% mutate(src = factor(src, levels = c(
                                      "regression_freq", 
                                      "regression_freq_distinct",
                                      "some_model",
                                      "habituals_model"
                                     ),
                      labels = c(
                                 "Referent frequency", 
                                 "Distinctiveness + \n referent frequency",
                                 "Fixed semantics",
                                 "Uncertain semantics"
                                 ) )) %>%
ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.15)+
  geom_errorbarh(alpha = 0.15)+
  geom_point(shape = 21, size = 3)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human endorsement")+
  facet_wrap(~src, nrow = 1, strip.position = "top")+
  guides(fill = guide_colorbar(title = 'log annual rate',
                               ticks = F,
                               title.position="top"))+ 
  theme(strip.text.y = element_blank(),# element_text(angle = 0, hjust= 0),
        legend.direction = 'horizontal',
        #legend.position = c(0.025,0.975),
        #legend.position =c(1,1),
        legend.justification = c(1,1)
        )
habituals.endorsement.models
```

@@

While past frequency is often a good indicator of future tendency, people change abruptly due to a variety of decisions and outside events.
Does habitual language communicate probabilities in terms of past frequency or future expectations?
On one hand, speakers can only be certain about what has happened in the past.
On the other hand, language has a communicative function, and it would seem useful for speakers to convey their predictions of what will be the case in the future.

We address this by introducing causal events that either enable or prevent future actions.
In one condition, participants make a prediction about what will be the case in the future (*predictive frequency*).
In another condition, participants decide whether or not to endorse the habitual sentence (as in Expt. 2b).
<!-- In Expt. 2d, we examine endorsements of the habitual sentence (e.g., \emph{John smokes cigarettes.}; \emph{Susan eats peanut butter.}) in the presence of these causal modifiers. -->
We then compare two uncertain-thresholds models to each other: one which uses participants' ratings of *predictive frequency* as the object of communication and one which uses the *past frequency* (as was done for Expt. 2b).

@@

@@

The events used were a subset of those used in Expts. 2a \& b (21 of the original 31).
In addition, we crafted statements that were intended to either increase the frequency (*enabling*; e.g., "Yesterday, Bill bought a pack of cigarettes.") or decrease the frequency (*preventative*; "Yesterday, Bill quit smoking.") of the event in the future.
In order to increase the potential variability of responses across the experimental conditions, participants only saw the frequencies that led to the most intermediate endorsement of the habitual in Expt. 2b.
We did not include separate trials for both male and female names for the select items we did in Expt. 2b, since we saw no differences in their endorsements of the habitual.
See Table \mht{one table for Expt 2 stimuli} for a full list of the enabling and preventative information.

@@

Figure \ref{fig:habituals-predictive-figure} (left) shows the mean predicted future frequency as a function of the past frequency given to the participant and the type of causal information given. 
We observe in the baseline condition that future frequency perfectly tracks past frequency. 
That is, participants believe if a person smoked cigarettes 3 times last month, they will smoke cigarettes 3 times next month. 
This result implies that our model makes identical predictions for Expt. 2b whether the target is past frequency or expected future frequency (indicating, as expected, that we must look to the new data to distinguish these models).
Critically, we observe the preventative information strongly decreases and the enabling information slightly increases predicted frequency (Figure \ref{fig:habituals-predictive-elicitation} left; blue and green dots).

@@

We use our formal model to test whether past or predictive frequency matters for endorsement.
To formalize the predictive frequency speaker model, we use the mean predictive frequency as the referent-prevalence $h'$ that the endorsement model (Eq. \ref{eq:S1}) conditions on. 
The past frequency model is constructed using the past frequency supplied to participants as the referent-prevalence.
We analyze this model in the same Bayesian data analysis regime as for our previous models.
We use the same priors over the parameters as before, and learn about the posterior distribution by collecting three independent MCMC chains of 100,000 iterations (removing the first 50,000 for burn-in). 
Figure \ref{fig:habituals-predictive-figure}C shows the resulting model predictions for the past frequency and the predictive frequency endorsement models. 
Participants' judgments of the habitual statements was indeed influenced by the causal manipulations *in the way predicted* by the endorsement model that uses the predictive frequency as the object of communication ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.predictive`$; MSE = $`r mse.habituals.rsa.predictive`$).
The model based on past frequency does not make different predictions for the different causal manipulation conditions and does a poor job at explaining the endorsements ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.past`$; MSE = $`r mse.habituals.rsa.past`$).

>>>>

```{r habituals-endorsement-gender-exploration-scatter, fig.cap='Exploratory analysis of endorsements of habitual statements by gender of target character (e.g., "Mary drinks beer" vs. "John...") for six events that displayed appreciable differences in the prior elicitation. Labeled points all correspond to the time period of 3 times / month.'}
gendered.items <- c("does cocaine", "drinks beer", "drinks coffee",
                    "wears a bra", "wears a watch", "wears a suit")

@@

hab.uncertain.threshold.s1opt <- m.hab.samp %>% 
  filter(type == "param") %>%
  group_by(B) %>%
  summarize(MAP = estimate_mode(val),
            cred_upper = hdi_upper(val),
            cred_lower = hdi_lower(val))

@@

habituals.endorsement.models <- md.hab %>% mutate(src = factor(src, levels = c(
                                      "regression_freq", 
                                      "regression_freq_distinct",
                                      "some_model",
                                      "habituals_model"
                                     ),
                      labels = c(
                                 "Referent frequency", 
                                 "Distinctiveness + \n referent frequency",
                                 "Fixed semantics",
                                 "Uncertain semantics"
                                 ) )) %>%
ggplot(., aes ( x = prediction, xmin = prediction_lower, xmax = prediction_upper,
                  y = MAP_h, ymin = low, ymax = high, fill = logAnnualRate))+
  geom_abline(intercept = 0, slope = 1, lty = 3)+
  geom_linerange(alpha = 0.15)+
  geom_errorbarh(alpha = 0.15)+
  geom_point(shape = 21, size = 3)+
  scale_x_continuous(limits = c(-0.01, 1.01), breaks = c(0,  1))+
  scale_y_continuous(limits = c(-0.01, 1.01), breaks = c(0, 1))+
  #scale_fill_continuous(low = "#2b83ba", high = "#d7191c")+
  coord_fixed()+
  xlab("Model prediction")+
  ylab("Human endorsement")+
  facet_wrap(~src, nrow = 1, strip.position = "top")+
  guides(fill = guide_colorbar(title = 'log annual rate',
                               ticks = F,
                               title.position="top"))+ 
  theme(strip.text.y = element_blank(),# element_text(angle = 0, hjust= 0),
        legend.direction = 'horizontal',
        #legend.position = c(0.025,0.975),
        #legend.position =c(1,1),
        legend.justification = c(1,1)
        )
```

@@

While past frequency is often a good indicator of future tendency, the future is under no obligation to mimic the past.
Does habitual language communicate probabilities in terms of past frequency or future expectations?
On one hand, speakers can only be certain about what has happened in the past.
On the other hand, language has a communicative function, and it would be useful for speakers to convey their predictions of what will be the case in the future.

People can change their behavior abruptly due to a variety of decisions and outside events.
We introduce these causal events into our minimal endorsement paradigm and measure their influence on endorsement.
<!-- that either enable or prevent future actions. -->
To provide the appropriate model-based analysis, participants in one condition make a prediction about what will be the case in the future (*predictive frequency*).
In another condition, participants decide whether or not to endorse the habitual sentence (as in Expt. 2b).
<!-- In Expt. 2d, we examine endorsements of the habitual sentence (e.g., \emph{John smokes cigarettes.}; \emph{Susan eats peanut butter.}) in the presence of these causal modifiers. -->
We then compare two uncertain-thresholds models to each other: one which uses participants' ratings of *predictive frequency* as the object of communication and one which uses the *past frequency* (as was done for Expt. 2b).

@@

r.hab.pred.past.baseline.freq <- round(with(d.hab.predictive.summary %>% filter(condition == "baseline"), cor(logAnnualPastRate, mean)), 3)
n.hab.pred.past.baseline.freq <- length(filter(d.hab.predictive.summary, condition == "baseline")$mean)

@@

The events used were a subset of those used in Expts. 2a \& b (21 of the original 31).
In addition, we crafted statements that were intended to either increase the frequency (*enabling*; e.g., "Yesterday, Bill bought a pack of cigarettes.") or decrease the frequency (*preventative*; "Yesterday, Bill quit smoking.") of the event in the future.
In order to increase the potential variability of responses across the experimental conditions, participants only saw the frequencies that led to the most intermediate endorsement of the habitual in Expt. 2b.
We did not include separate trials for both male and female names for the select items we did in Expt. 2b, since we saw no differences in their endorsements of the habitual.
See Appendix D for a full list of the enabling and preventative information.

@@

Figure \ref{fig:habituals-predictive-figure} (left) shows the mean predicted future rate as a function of the past rate given to the participant and the type of causal information given. 
We observe in the baseline condition that future frequency perfectly tracks past frequency ($r(`r n.hab.pred.past.baseline.freq`) = `r r.hab.pred.past.baseline.freq`$). 
That is, participants believe if a person smoked cigarettes 3 times last month, they will smoke cigarettes 3 times next month. 
This result implies that our model makes identical predictions for Expt. 2b whether the referent is past frequency or expected future frequency (indicating, as expected, that we must look to the new data to distinguish these models).
Critically, we observe the preventative information strongly decreases and the enabling information slightly increases predicted frequency (Figure \ref{fig:habituals-predictive-elicitation} left; blue and green dots).

@@

We use our formal model to test whether past or predictive frequency matters for endorsement.
To formalize the predictive frequency speaker model, we use the mean predictive frequency as the referent-prevalence $p_{fk}$ that the endorsement model (Eq. \ref{eq:S1}) conditions on. 
The past frequency model is constructed using the past frequency supplied to participants as the referent-prevalence.
We analyze this model in the same Bayesian data analysis regime as for our previous models.
We use the same priors over the parameters as before and learn about the posterior distribution by collecting three independent MCMC chains of 100,000 iterations (removing the first 50,000 for burn-in). 
Figure \ref{fig:habituals-predictive-figure}C shows the resulting model predictions for the past frequency and the predictive frequency endorsement models. 
Participants' judgments of the habitual statements was indeed influenced by the causal manipulations *in the way predicted* by the endorsement model that uses the predictive frequency as the object of communication ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.predictive`$; MSE = $`r mse.habituals.rsa.predictive`$).
The model based on past frequency does not make different predictions for the different causal manipulation conditions and does a poor job at explaining the endorsements ($r^2(`r r2.hab.rsa.predictive.n`) = `r r2.habituals.rsa.past`$; MSE = $`r mse.habituals.rsa.past`$).
